---
title: "Visualization 3: Group by and Summarize"
output: github_document
---

```{r, message = FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(viridis)
```

Load in a data set that we will use often (copied from p8105 website).

```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```

## `group_by`

```{r}
weather_df %>%
  group_by(name, month)
```


## counting things

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(n_obs = n())

weather_df %>% 
  count(name)
```

its nice that summarize produces a data frame. 

```{r}
weather_df %>% 
  pull(name) %>% 
  table()
```

Using `table` does NOT give us a data frame, If we try to use this in ggplot it breaks. 

```{r}
weather_df %>% 
  janitor::tabyl(month, name)
```

this is maybe useful if we need to make a table for someone else to read. 




## MOre general summaries

Let's count, take means, looks at SDs

```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(
    n_obs = n(),
    mean_tmax = mean(tmax, na.rm = T),
    median_tmax = median(tmax, na.rm = T),
    sd_tmax = sd(tmax, na.rm = T)
  )
```

adding additional groups and graphing

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    n_obs = n(),
    mean_tmax = mean(tmax, na.rm = T),
    median_tmax = median(tmax, na.rm = T),
    sd_tmax = sd(tmax, na.rm = T)
  ) %>% 
  ggplot(aes(x = mean_tmax, y = month, color = name)) +
  geom_point() +
  geom_line()
```

Formatting table outputs

```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(
    n_obs = n(),
    mean_tmax = mean(tmax, na.rm = T)
  ) %>% 
  knitr::kable(digits = 2)
```

## grouped `mutate`

```{r}
weather_df %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = T)
  )
```

gives mean for all observations

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = T),
    centered_tmax = tmax - mean_tmax
  ) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) +
  geom_point()
```

Gives mean separate for each station (we also made a new centered variable and plotted it)


```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    tmax_rank = min_rank(tmax)
  ) %>% 
  filter(tmax_rank < 2)
```

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    tmax_rank = min_rank(desc(tmax))
  ) %>% 
  filter(tmax_rank < 2)
```

Lagged variables

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    lagged_tmax = lag(tmax, n = 1),
    tmax_dif = tmax - lagged_tmax
  ) 
```

using summarize to find the sd of `tmax-dif`

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    lagged_tmax = lag(tmax, n = 1),
    tmax_dif = tmax - lagged_tmax
  ) %>% 
  summarize(dif_sd = sd(tmax_dif, na.rm = T))
```


## Limitations

what if my "summary" is a linear model

```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(cor_tmin_tmax = cor(tmin, tmax, use = "complete"))

weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  lm(tmax ~ tmin, data = .)
```

