---
title: "Visualization 2: ggplot 2"
output: github_document
---

```{r, message = FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(viridis)
```

Load in a data set that we will use often (copied from p8105 website).

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Starting with a familiar one

```{r}
weather_df %>%  
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three sations",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package"
  )
```

## Scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three sations",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package") +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15") +
  scale_y_continuous(
    trans = "sqrt", 
    position = "right"
  ))
```

Hues

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three sations",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) +
  scale_color_hue(
    name = "Location",
    h = c(100, 300)
  ) +
  scale_color_viridis_d()
```


## Themes

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three sations",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package with three stations"
  ) +
  scale_color_viridis_d() +
  theme_bw() +
  theme(legend.position = "bottom")
```

for themes, order MATTERS. For `theme_bw()` it applies ALL defaults for the black and white theme, which means it would overwrite any themes listed before it. This is why we put the `theme(legend.position = "bottom")` command after the `theme_bw()`

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three sations",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package with three stations"
  ) +
  scale_color_viridis_d() +
  theme_minimal() +
  theme(legend.position = "bottom")
```

It is possible to set you themes as a global setting! Put the following code chunk at the **START** of an RMD

```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## `data` in geoms

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()

waikiki %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park)
```


## `patchwork`

```{r}
ggp_tmax_tmin = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  theme(legend.position = "none")

gg_prcp_dens = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.3) +
  theme(legend.position = "none")

ggp_tmax_date = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth() +
  theme(legend.position = "bottom")

(ggp_tmax_tmin + gg_prcp_dens) / ggp_tmax_date
```

The lowest line of code utilizes `patchwork` to combine the plots together!! It is NOT intuitive; do not forget this is how it works. You must have the library loaded to do this. 


## Data Manipulation

quick examples on factors

```{r}
weather_df %>% 
  mutate(
    name = fct_reorder(name, tmax)
  ) %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```


What about tmax and tmin . . . 

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "obs",
    values_to = "temperature"
  ) %>% 
  ggplot(aes(x = temperature, fill = obs)) +
  geom_density(alpha = .3) +
  facet_grid(. ~ name)
```


%
```{r}
pulse_df = 
  haven::read_sas("data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi",
    names_prefix = "bdi_score_"
  ) %>% 
  mutate(visit = recode(visit, "bl" = "00m"))

pulse_df %>% 
  ggplot(aes(x = visit, y = bdi)) +
  geom_boxplot()

pulse_df %>% 
  ggplot(aes(x = visit, y = bdi)) +
  geom_boxplot() +
  geom_line(aes(group = id))
```

Look at the course website for the last example!
